---
title: "Day 6 Activity"
date: "21 August 2023"
format: html
editor: visual
execute: 
  warning: false
---

## *Data Organization in Spreadsheets*

- people are often not consistent with their naming
- choose clear names that are easy for a computer to read as well
- YYYY-MM-DD is a good standard
- don't have empty cells- put NA (and be consistent with how you note NAs)
- don't do calculations in the raw excel file


## SBC Lobsters

```{r}
library(tidyverse)
library(here)
```

```{r}
lobster <- read_csv(here("data", "Lobster_Abundance_All_Years_20210412.csv"), na = c("-99999", "")) %>%
  janitor::clean_names() %>%
  uncount(count)
```

```{r}
mean_count <- lobster %>%
  group_by(site, year) %>%
  summarise(mean = mean(size_mm),
            count = n())
```

```{r}
ggplot(mean_count, aes(year, count, fill = site)) +
  geom_bar(stat = "Identity") +
  facet_wrap(~site) +
  theme_linedraw() +
  theme(legend.position = "none")
```

#### Legal Lobsters

```{r}
lobster_2020 <- lobster %>%
  filter(year == 2020)

lobster_2020_legal <- lobster_2020 %>%
  group_by(site) %>%
  mutate(legal = case_when(size_mm > 79.76 ~ "yes",
                           size_mm <= 79.76 ~ "no")) %>%
  count(legal)
```

```{r}
ggplot(lobster_2020_legal, aes(site, n, fill = legal)) +
  geom_col(position = "fill") +
  theme_linedraw()
```

#### Lobster Wrangling

```{r}
ex_a <- lobster %>%
  filter(site %in% c("IVEE", "CARP", "NAPL"))

ex_b <- lobster %>%
  filter(month == 8)

ex_c <- lobster %>%
  filter(site == "AQUE" | size_mm > 70)

ex_d <- lobster %>%
  filter(site != "NAPL")
```

#### group_by() and summarise() practice

```{r}
lobster %>%
  group_by(site) %>%
  summarise(mean = mean(size_mm, na.rm = TRUE),
            sd = sd(size_mm, na.rm = TRUE))

lobster %>%
  group_by(site, month) %>%
  summarise(max_length = max(size_mm, na.rm = TRUE))
```

#### mutate()

```{r}
size_conv <- lobster %>%
  mutate(size_cm = size_mm / 10)

lobster %>%
  mutate(site = tolower(site))

lobster %>%
  mutate(area = as.character(area))
```

#### case_when()

```{r}
size_bin <- lobster %>%
  mutate(size_bin = case_when(size_mm <= 70 ~ "small",
                              size_mm > 70 ~ "large"))
size_bin

lobster %>%
  mutate(designation = case_when(site == "IVEE" | site == "NAPL" ~ "MPA",
                                 TRUE ~ "not MPA"))
```




